### 2014-11-13 tex/intro.tex
Rewrote some portions on the introducation based on Jay's feedback

Initial work on symbolic execution largely focused on checking 
properties of programs with primitive types~\cite{clarke76TSE,King:76}.
%, e.g., integers and booleans 
%Despite recent advances
%in constraint solving technologies, improvements in
%raw computing power, and advances in reduction
%and abstraction techniques~\cite{AnandETAL2009AbsSymExe,Godefroid:POPL07}
%symbolic execution of programs of modest size containing only
%primitive types, remains challenging
%because of the large number of execution paths generated
%during symbolic analysis. 
With the advent of object-oriented languages,
% e.g., Java and C$++$, 
recent work has generalized the core ideas of symbolic execution to enable 
analysis of programs containing complex data structures with unbounded 
domains, i.e., data stored on the heap~\cite{Kiasan06,Kiasan07,GSE03}.  
These Generalized Symbolic Execution (GSE) techniques construct the 
heap in a lazy manner, deferring
materialization of objects on the concrete heap until they are needed
for the analysis to proceed. The materilization 
%accounts for various heap configurations, 
creates additional non-deterministic choice points in the symbolic
execution tree representing the feasible heap configurations.  as (a)
null, (b) instance of a new reference of a compatible class, and (c)
alias an previously initialized symbolic reference.  Symbolic
execution follows concrete program semantics for materialized heap
locations. Although this approach enables analysis of heap
manipulating programs, a large number of feasible concrete heap
configurations may be created exacerbating the well known state space
explosion problem associated with symbolic execution.

@@

The goal of this work is to mitigate the state space explosion problem
in GSE by grouping multiple heaps together and only partitioning the
heaps at points of divergence in the control flow graph. To achieve
this goal, we found inspiration from recent work in the domain of
static analysis that uses sets of constraints over heap locations to
encode multiple heaps in a single representation. These sets,
sometimes known as value sets or value summaries, allow multiple heaps
to be represented simultaneously with a higher degree of precision
than afforded by traditional shape analysis techniques. Some of these
previous attempts, however, are unable to handle aliasing in heaps due
to a recursive definition of objects~\cite{..}, while others require a
set of heaps as input and are unable to initialize heaps
on-the-fly~\cite{..}. For these reasons the techniques use
over-approximations to reason about heaps and may report false
positives. Also, update to heaps require a rewrite of the constraint
system making it prohitive for use in the context of symbolic
execution.

@@

In this work, we present a technique for heap summarization,
initialization, and analysis in the context of symbolic execution. We
represent the heap as a bipartiate graph of references and sets of
constraints over locations. The use of a bipartiate graph affords two
key advantages: (i) it allows for a non-recursive definition of
objects, allowing us to handle aliasing and (ii) it does not require
rewrite of the constraints during updates to the heap nor does it
introduce any auxillary variables in the constraints. During symbolic
execution, we initialize, summarize, and update the bipartiate
graph. Unlike GSE, however, the heaps are partitioned only at points
of divergence in the control flow graph. These points of
non-determinism are at field accesses or writes that lead to
null-pointer exceptions and at comparisons of references. The former
represents a divergence due to exceptional control flow while the
later due to program structure.

@@

The heap summary approach for symbolic execution is sound and complete
with respect to properties that are provable using GSE. This is
accomplished by showing the existence of a bisimulation relation. In
other words, any property that can be proved using GSE can also be
proved using heap summaries, furthermore, any property proved by heap
summaries can also be proved by GSE---given the same bounds on heap
reference chains and loops.

In a preliminary evaluation comparing the efficiacy of the heap
summary technique with respect to the other GSE approaches, we observe
a two-order of magnitude reduction in the total time taken to explore
the same state space. We implement the heap summary approach as an
extension to the Symbolic PathFinder symbolic execution engine and
analyze three Java data structure artifacts. For these examples, we
show that where other GSE approaches are unable to complete
exploration within the provided time bound, the summary heap finishes
exploration in a few seconds.

@@

\item\textbf{-} Define a heap encoding as a bipartiate graph of
  references and sets of constraints over locations.

@@

\item\textbf{-} Present an algorithm to initialize, summarize, and
  update the bipartiate graphs during symbolic execution.

\item\textbf{-} Show the existence of a bisimulation relation prove
  that the heap summary approach is sound and complete with respect to
  properties provable by GSE.

\item\textbf{-} Demonstrate the scability of the heap summary approach
  compared to GSE approaches.

@@

\noindent{In this paper, we present a brief over of the a Java-like
  language and the encoding of heaps as a bipartiate graph. We use the
  Java-like language to define the formal semantics of the heap
  summary machine and then prove the existence of a bisumulation
  relation. Finally we present an evaluation of the heap summary
  approach. }

>>>>

Initial work on symbolic execution largely focused on checking 
properties of programs with primitive types~\cite{clarke76TSE,King:76}.
%, e.g., integers and booleans 
%Despite recent advances
%in constraint solving technologies, improvements in
%raw computing power, and advances in reduction
%and abstraction techniques~\cite{AnandETAL2009AbsSymExe,Godefroid:POPL07}
%symbolic execution of programs of modest size containing only
%primitive types, remains challenging
%because of the large number of execution paths generated
%during symbolic analysis. 
With the advent of object-oriented languages,
% e.g., Java and C$++$, 
recent work has generalized the core ideas of symbolic execution to enable 
analysis of programs containing complex data structures with unbounded 
domains, i.e., data stored on the heap~\cite{Kiasan06,Kiasan07,GSE03}.  
These \emph{Generalized Symbolic Execution} (GSE) techniques construct the 
heap in a lazy manner, deferring
materialization of objects on the concrete heap until they are needed
for the analysis to proceed. The materilization 
%accounts for various heap configurations, 
creates additional non-deterministic choice points in the symbolic
execution tree by representing the feasible heap configurations as (i)
null, (ii) an instance of a new reference of a compatible class, and (iii)
an alias to a previously initialized symbolic reference.  Symbolic
execution then follows concrete program semantics for materialized heap
locations. Although this approach enables the analysis of heap
manipulating programs, a large number of feasible concrete heap
configurations are created leading to state space explosion.

@@

The goal of this work is to mitigate the state space explosion problem
in GSE by grouping multiple heaps together and only partitioning the
heaps at points of divergence in the control flow graph. Our
inspiration is found in the domain of static analysis that uses sets
of constraints over heap locations to encode multiple heaps in a
single representation. These sets, sometimes known as value sets or
value summaries, allow multiple heaps to be represented simultaneously
with a higher degree of precision than afforded by traditional
techniques for shape analysis.

Unfortunately, value sets often do not support aliasing and require a
recursive definition of objects~\cite{..}. Also, it is not obvious how
to support heap updates in value sets, and in particular, on-the-fly
lazy initialization as is common in GSE, so it is not unusual to
require a fully initialized set of input heaps in the analysis. As is
typical in static analysis, over approximation of the value sets
alleviates some of the limitations but also leads to false positives,
and when heap updates are supported, with or without over
approximation, the path constraint must be rewritten to reflect the
update, and the constraints in the value set must also be rewritten.

@@

The solution presented in this paper to enabling value sets in symbolic execution
with lazy initialization is to represent the heap as a bipartite graph
consisting of references and locations. A reference is able to point
to several different locations; each edge is predicated on constraints
over aliasing between references. Locations point back to references
on different fields. The bipartite graph enables (i) the use of
aliasing to define objects rather than recursive definitions; (ii)
value set computation using only local information; and (iii) heap
updates that only impact affected locations, require no constraint
rewriting, and require no auxiliary variables in the constraints.  We
call this new heap a summary heap.

To mitigate state space explosion in GSE, the solution described in
paper uses a summary heap in the symbolic execution. This summary heap
algorithm defines how the bipartite graph updates in lazy
initialization. Unlike GSE, however, the summary heap algorithm
partitions only at points of divergence in the control flow
graph. These points of non-determinism are at field accesses or writes
that lead to null-pointer exceptions and at comparisons of
references. The former represents a divergence due to exceptional
control flow while the later due to program structure.

@@

Importantly, the summary heap algorithm is sound and complete with
respect to properties that are provable using GSE. This proof is
accomplished by showing the existence of a bisimulation relation
between states in GSE and states in the summary heap algorithm. In
practical terms, any property that can be proved using GSE can also be
proved using summary heaps, furthermore, any property proved by
summary heaps can also be proved by GSE. Of course, Both require a bound on heap
reference chains and loops.

Aside from the novelty of the heap summary and it application to
symbolic execution, the new data structure appears to ameliorate state
explosion in symbolic execution. A preliminary implementation in Java
Pathfinder (JPF), described in this paper, compares the heap summary
algorithm on a handful of benchmarks, each representing different
degrees of initialization and branching, to GSE with lazy
initialization in Symbolic Pathfinder (SPF) and implementation
lazier\# also in SPF. The results show that in general, that the heap
summary algorithm improves over other techniques, and in some
instances, the improvement is remarkable: two-orders of magnitude
reduction in running time. More critically though, the heap summary
enables the initialization of larger more complex heaps than
heretofore possible as shown in our results.

\begin{comment}
We implement the summary heap algorithm as an extension to \emph{Java
  PathFinder} (JPF) and compare it to GSE with lazy initialization in
\emph{Symbolic Pathfinder (SPF)} and an implementation of lazier\#
initialization in SPF.  The comparison shows on some examples up to a two-order of magnitude
reduction in the total time taken to explore the same state space
defined by the bound on the longest heap reference chain.  For these
examples, we show that where other GSE approaches are unable to
complete exploration within the provided time bound, the summary heap
finishes exploration in a few seconds.
\end{comment}

@@

\item\textbf{-} A summary heap as a bipartite graph that supports aliasing, local value set computation, and heap updates.

@@

\item\textbf{-} A summary heap algorithm to initialize, summarize, and
  update the bipartiate graphs during symbolic execution.

\item\textbf{-} A proof of the existence of a bisimulation relation to establish
  that the heap summary approach is sound and complete with respect to
  properties provable by GSE.

\item\textbf{-} A proof of concept implementation and small empirical study to demonstrate the scability of the heap summary approach
  compared to other GSE approaches.

@@

\noindent{The rest of this paper discusses a formal language semantics
  for the summary heap algorithm definition with the definition of the
  summary heap. The paper then presents the application of summary
  heaps to symbolic execution. Following the algorithm definition, the
  existence proof of bisimulation is given with the accompanying
  corollaries for soundness and completeness. This is followed by the
  implementation in JPF and the evaluation. The paper ends with the
  conclusion and future work.}

### 2014-11-13 tex/precise.tex
Rewrote some portions on the introducation based on Jay's feedback

The rewrite rule to initialize a heap in the summary heap algorithm
is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $\cfgnt{r}_f$,
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference-location pairs, $(\cfgnt{r}_a\ \cfgnt{l}_a)$, the
field being initialized may alias. A pair $(\cfgnt{r}_a\ \cfgnt{l}_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $\cfgnt{r}_a$ is
part of the input references initialized in the summary heap, (ii)
$\cfgnt{r}_a$ is in the pre-image of $\cfgnt{l}_a$, and (iii) the type 
of $\cfgnt{l}_a$ is the same as the type of the field being initialized.

>>>>

The rewrite rule to initialize a heap in the summary heap algorithm
is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $\cfgnt{r}_f$,
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference-location pairs, $(\cfgnt{r}_a\ \cfgnt{l}_a)$, the
field being initialized may alias. A pair $(\cfgnt{r}_a\ \cfgnt{l}_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $\cfgnt{r}_a$ is
part of the input references initialized in the summary heap, (ii)
$\cfgnt{r}_a$ is in the pre-image of $\cfgnt{l}_a$, and (iii) the type 
of $\cfgnt{l}_a$ is the same as the type of the field being initialized.
